\bf Are you normal? A new projection pursuit index to assess a sample against a multivariate null distribution

### 2024-06-27 paper.tex

The projection of an ellipsoid onto \gD{} is an ellipse, where the curve
of the ellipse is defined through the set of points \(\mathbfit{x}\) for
which the gradient is parallel to the projection plane. We call points
in the projection that are on the curve \(\mathbfit{y}\). From this we
can compute the analog equation for the projection as

\[(\mathbfit{y} - \mathbfit{\mu}_p)(P^T \Sigma P)^{-1}(\mathbfit{y} - \mathbfit{\mu}_p)^T = c^2\]
where \(P\) is a \((p\times 2)\) orthonormal basis defining the
projection and \(\mathbfit{\mu}_p = \mathbfit{\mu} P\) the projected
mean. This means the matrix \((P^T \Sigma P)^{-1}\) is defining the
ellipse in the \gD{} projection. In general \(c\) could be any constant,
but typically we would select it as a quantile of the \(\chi^2\)
distribution, so that the size of the ellipse corresponds to a selected
probability.

>>>>

The projection of an ellipsoid onto \gD{} is an ellipse, where the curve
of the ellipse is defined through the set of points \(\mathbfit{x}\) for
which the gradient is parallel to the projection plane. That is, the curve 
consists of \(\mathbfit{x}\) satisfying
\[
\nabla (\mathbfit{x}-\mathbfit{\mu}) \Sigma^{-1}(\mathbfit{x}-\mathbfit{\mu})^T = 2 (\mathbfit{x}-\mathbfit{\mu}) \Sigma^{-1} = 2\mathbfit{s}P^T
\]
for some \(\mathbfit{s} \in \mathbb{R}^2\), where \(P\) is a \((p\times 2)\) orthonormal basis defining the
projection. We can write \(\mathbfit{x}-\mathbfit{\mu} = \mathbfit{s} P^T \Sigma\).
Making this substitution in the \pD{} ellipsoid equation yields
\[
c^2 = (\mathbfit{x}-\mathbfit{\mu}) \Sigma^{-1}(\mathbfit{x}-\mathbfit{\mu})^T = \mathbfit{s} P^T \Sigma P \mathbfit{s}^T
\]

We call points in the projection that are on the curve \(\mathbfit{y}\), so that \(\mathbfit{y} = \mathbfit{x}P\)
for \(\mathbfit{x}\) on the \pD{} curve. Then
\(\mathbfit{y} - \mathbfit{\mu}_p = (\mathbfit{x} - \mathbfit{\mu})P = \mathbfit{s} P^T \Sigma P\), where 
\(\mathbfit{\mu}_p = \mathbfit{\mu} P\) is the projected mean. We then
substitute \((\mathbfit{y} - \mathbfit{\mu}_p) (P^T \Sigma P)^{-1}\) for \(\mathbfit{s}\) in the
previous equation. From this we
can compute the analog equation for the projection as

\[(\mathbfit{y} - \mathbfit{\mu}_p)(P^T \Sigma P)^{-1}(\mathbfit{y} - \mathbfit{\mu}_p)^T = c^2\]

This means the matrix \((P^T \Sigma P)^{-1}\) is defining the
ellipse in the \gD{} projection. In general \(c\) could be any constant,
but typically we would select it as a quantile of the \(\chi^2\)
distribution, so that the size of the ellipse corresponds to a selected
probability.
